---
title: "Getting Started with statsforensics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with statsforensics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(statsforensics)
```

## Overview

`statsforensics` provides a cohesive set of functions for the full statistical
analysis cycle: exploratory data analysis, assumption checking, hypothesis
testing, regression modeling, and diagnostics. This vignette walks through a
complete workflow using the built-in `vehicles` dataset.

---

## 1. The Vehicles Dataset

The package ships with `vehicles`, a cleaned subset of the EPA fuel economy
data derived from `ggplot2::mpg`.

```{r data}
data(vehicles)
head(vehicles)
```

Check for missing values before doing anything else:

```{r missing}
count_missing(vehicles)
```

---

## 2. Exploratory Data Analysis

### Single variable

`eda_numeric()` produces a histogram, boxplot, and Q-Q plot alongside summary
statistics. Set `quiet = FALSE` to print the statistics to the console.

```{r eda-numeric}
eda_numeric(vehicles, "highway_mpg", quiet = FALSE)
```

### Comparing groups

`eda_by_group()` overlays boxplots and per-group Q-Q plots to reveal distributional
differences before any formal test.

```{r eda-by-group}
eda_by_group(vehicles, "highway_mpg", "vehicle_class")
```

### Group-wise summary table

```{r summarize-group}
summarize_by_group(vehicles, c("highway_mpg", "city_mpg"), "vehicle_class")
```

---

## 3. Assumption Checking

### Normality

`check_normality()` runs Shapiro-Wilk and Kolmogorov-Smirnov tests and reports
an interpretation.

```{r normality}
check_normality(vehicles$highway_mpg)
```

### Homogeneity of variance

Before running ANOVA, verify that variance is roughly equal across groups.

```{r levene}
check_heteroscedasticity(vehicles, response = "highway_mpg", group = "vehicle_class")
```

---

## 4. Hypothesis Testing

### One-way ANOVA

```{r anova}
anova_result <- perform_anova(vehicles, response = "highway_mpg", group = "vehicle_class")
```

The effect size η² shows how much of the variance in highway MPG is explained
by vehicle class.

### Post-hoc comparisons

When ANOVA is significant, Tukey HSD identifies which pairs differ.

```{r tukey}
tukey_post_hoc(anova_result)
```

### Non-parametric alternative

If normality is violated, use the Kruskal-Wallis test instead.

```{r kruskal}
kruskal_wallis_test(vehicles, response = "highway_mpg", group = "vehicle_class")
```

---

## 5. Regression Modeling

### Fit a linear model

```{r lm}
model <- fit_linear_model(vehicles, "highway_mpg ~ engine_displacement")
```

### Model summary

`model_summary()` presents fit statistics and coefficient table in a clean format.

```{r model-summary}
model_summary(model)
```

### Validate assumptions

After fitting, check whether the model's residuals satisfy the regression
assumptions before drawing inference.

```{r validate}
validate_assumptions(model)
```

### Diagnostic plots

`plot_diagnostics()` generates a 4-panel grid: residuals vs fitted, normal Q-Q,
scale-location, and residual order plot.

```{r diagnostics}
plot_diagnostics(model)
```

---

## 6. Predictions

Generate predictions with uncertainty intervals on a new grid of values.

```{r predictions}
new_data <- data.frame(engine_displacement = seq(1.6, 7.0, by = 0.5))
preds <- create_prediction_intervals(model, newdata = new_data)
plot_predictions(preds)
```

---

## 7. Comparing Models

```{r compare}
model_simple <- fit_linear_model(vehicles, "highway_mpg ~ engine_displacement")
model_full   <- fit_linear_model(vehicles, "highway_mpg ~ engine_displacement + cylinders")

compare_models(model_simple, model_full)
```

Use `model_dashboard()` to compare many models at once on key metrics.

```{r dashboard}
model_poly <- fit_polynomial_model(vehicles, "highway_mpg", "engine_displacement", degree = 2)

model_dashboard(list(
  Linear      = model_simple,
  Full        = model_full,
  Polynomial  = model_poly
))
```

---

## Next Steps

- See `?check_multicollinearity` for VIF-based collinearity diagnosis on
  multi-predictor models.
- Use `chi_squared_test()` and `correspondence_analysis_wrapper()` for
  categorical association analysis.
- `generate_report()` wraps the entire pipeline into a single console report.
